Data-Driven Search-Based Software Engineering

### 2018-01-17 main.tex
Over to Dr. M.

@@

@@

\begin{document}
\title{Research and Teaching
Tools for Data-Driven Search-Based Software Engineering}
% \titlenote{Produces the permission block, and
%   copyright information}
% \subtitle{Extended Abstract}
% \subtitlenote{The full version of the author's guide is available as
%   \texttt{acmart.pdf} document}

\author{Vivek Nair, everyone else, Tim Menzies}
% \authornote{Dr.~Trovato insisted his name be first.}
% \orcid{1234-5678-9012}
\affiliation{%
  \institution{North Carolina State University, USA}
%   \streetaddress{P.O. Box 1212}
  \city{Raleigh} 
  \state{NC} 
  \postcode{27606}
}
\email{dchen20@ncsu.edu,  wfu@ncsu.edu,  tim@menzies.us}

\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings.\footnote{This is an abstract footnote}
\end{abstract}

@@

\keywords{ACM proceedings, \LaTeX, text tagging}

@@

% \input{samplebody-conf}
\section{Introduction}

    \subsection{Why Actionable Software Analytics?}
        Software analytics aim to
        obtain actionable insights
        from software artifacts that
        help practitioners accomplish tasks 
        related to software development, systems, 
        and users.

        The key part of software analytics is to provide 
        actionable insights, which means it have to 
        ignore the unimportant and irrelevant information, 
        while provide the most important one clearly and directly.
        For example, if the manager is driving toward a cliff, 
        we don’t want to distract her with analytics 
        telling her about the clouds in the sky or 
        the flowers on the side of the road. Instead, 
        we want our smart analytics to shout in her ear, 
        “There’s a cliff up ahead! Turn left immediately!”

        So why actionable software analytics? 
        For software vendors, managers, developers and users, 
        the actionable insights provided by software analytics 
        are considered as the core deliverable~\cite{tan2016defining}. 
        In fact, for software analytics with no actionable advice provided, 
        it is hard to imagine its usefulness and existence. 
        More importantly,
        actionable insight is the key driver for businesses 
        to invest in data analytics initiatives~\cite{sawyer2013bi}.

    \subsection{Why heuristics work?}
        Heuristics, together with logic and probability 
        are three central ideas in the intellectual history of our mind.
        Unlike statistical optimization procedures, 
        heuristics do not try to find the best solution
        but rather find a good-enough solution). 
        Because of that, 
        people tends to believe heuristics produce second-best results, 
        while optimization is always better.
        In spite of this one, Gigerenzer pointed out the other five common but erroneous beliefs about heuristics\cite{gigerenzer2008heuristics}, e.g. "People rely on heuristics only in routine
        decisions of little importance", 
        "People with higher cognitive capacities employ
        complex weighting and integration of
        information; those with lesser capacities use
        simple heuristics", and 
        "More information and computation is always
        better".

        In the late 90s, Gigerenzer discovered 
        the take-the-best heuristic is
        both more frugal and more accurate than 
        multiple regression

        Czerlinski et al. confirmed 
        the take-the-best heuristic is
        both more frugal and more accurate than 
        multiple regression result for 20 real-world problems\cite{czerlinski1999good}. 
        In 2006, Brighton showed for the first time that 
        the take-the-best heuristic is 
        often more accurate and frugal than
        complex nonlinear algorithms, including neural networks, 
        exemplar models, and classification and regression trees~\cite{brighton2006robust}.

        So why do heuristics work? As Gigerenzer summarized in 2008, heuristics exploit evolved capacities that
        come for free, and thus they can provide solutions to problems
        that are different from strategies of logic and probability.

        The take-the-best heuristic is designed to help choose
        between two alternatives, thus could be adjusted for 
        classification tasks, where each object needs to be assigned to
        one of several classes. 
        Basically, the resulting classification rule is a 
        fast and frugal tree, which allows for a quick decision 
        at each node of the tree~\cite{martignon2003naive}.

    \subsection{What is Fast and Frugal Trees(FFTs)? }
        Fast-and-frugal trees were proposed by Martignon et al. at 2018
        as decision trees with exactly two branches extending
        from each node, where either one or both branches is an exit
        branch leading to a leaf~\city{martignon2008categorization}. 
        That is to say, in an FFT,  every question posed by a node will 
        trigger an immediate decision. Because FFTs have an exit
        branch on every node, they usually make decisions faster
        than common decision trees while
        simultaneously being easier to understand and use.

         provides an example of fast and frugal tree.

        ~\ref{fig:fft-demo} presents an FFT designed to classify a change
        as being bug or not.
        The three nodes in the FFT correspond to the results of
        three features/cues: wmc, rdc, ce

        % thal indicates the result of a thallium
        % scintigraphy, a nuclear imaging test that shows how well
        % blood flows into the heart while exercising or at rest. The
        % result of the test can either be normal (n), indicate a fixed
        % defect (fd), or a reversible defect (rd). The second node
        % uses the cue cp, indicating a patient’s type of chest pain,
        % which can be either typical angina (ta), atypical angina (aa),
        % non-anginal pain (np), or asymptomatic (a). Finally, ca
        % indicates the number of major vessels colored by fluoroscopy,
        % a continuous x-ray imaging tool, whose values can range
        % from 0 to 3.

        \begin{figure}[!t]
            \centering
            \includegraphics[width=3.1in]{img/fft-demo.png}
            \caption{An example Fast Frugal Tree on the log4j dataset. 
            (Generated by 
            \href{https://econpsychbasel.shinyapps.io/ShinyFFTrees/}{ShinyFFTrees} 
            from Nathaniel.)
            }
            \label{fig:fft-demo}
        \end{figure}

    \subsection{Why use FFTs?}
        A lot of recent works have shown that the fast and frugal tree 
        could performs comparably well to the complex models in different kinds of domains, e.g. public health, medical risk management, performance science, etc.~\cite{jenny2013simple, laskey2014comparing, raab2015power}

\section{Methods}
    \subsection{Data}

    \subsection{Models}

    \subsection{Evaluation}

\section{Results}

\section{Discussions}

\section{Conclusions}

>>>>

\usepackage{color}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{graphicx} 
\usepackage{multirow}
\usepackage{siunitx}
\usepackage{rotating}
\usepackage{eqparbox}
\usepackage{wrapfig}
\usepackage{mathtools}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{lstlinebgrd}
\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{lstautogobble}
\usepackage[framemethod=tikz]{mdframed}
\usetikzlibrary{shadows}
\newmdenv[tikzsetting= {fill=white!20},roundcorner=10pt, shadow=true]{myshadowbox}
\usepackage{graphics}
\usepackage{colortbl} 
\usepackage{multirow} 
\usepackage{balance}
\usepackage{picture}
\usepackage{soul}
\usepackage{array}
\usepackage{makecell}

\usepackage{times}
\usepackage{wasysym}

\renewcommand\theadalign{cb}
\renewcommand\theadfont{\bfseries}
\renewcommand\theadgape{\Gape[4pt]}
\renewcommand\cellgape{\Gape[4pt]}

\newcommand\floor[1]{\lfloor#1\rfloor}
\newcommand\ceil[1]{\lceil#1\rceil}
\newcommand{\bibemph}[1]{{\em#1}}
\newcommand{\bibemphic}[1]{{\em#1\/}}
\newcommand{\bibsc}[1]{{\sc#1}}

\usepackage{verbatim}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[export]{adjustbox}
\renewcommand{\footnotesize}{\scriptsize}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\usepackage{caption}

@@

\usepackage{verbatim}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

\definecolor{MyDarkBlue}{rgb}{0,0.08,0.45} 
\lstset{
    language=Python,
    basicstyle=\sffamily\fontsize{2.3mm}{0.8em}\selectfont,
    breaklines=true,
    prebreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
    frame=l,
    showtabs=false,
    escapechar=|,
    columns=fullflexible,
    showspaces=false,
    showstringspaces=false,
    keywordstyle=\bfseries\sffamily,
    emph={generate, fitness, recombination_op, elitism, ML.train, model, find_best_model, model_building}, emphstyle=\bfseries\color{blue!50!black},
    stringstyle=\color{green!50!black},
    commentstyle=\color{red!50!black}\it,
    numbers=left,
    captionpos=t,
}

@@

\begin{document}
\title{
% Research and Teaching Tools for 
Data-Driven Search-Based Software Engineering}
% \titlenote{Produces the permission block, and
%   copyright information}
% \subtitle{Extended Abstract}
% \subtitlenote{The full version of the author's guide is available as
%   \texttt{acmart.pdf} document}

\author{Vivek Nair, everyone else, Tim Menzies}
\affiliation{%
  \institution{North Carolina State University, USA}
  \city{Raleigh} 
  \state{NC} 
  \postcode{27606}
}
\email{}

\begin{abstract}
This paper introduces ideas and resources for a new field of software engineering research: Data-Driven Search-based software engineering (DSE), which includes insights from Software Analytics (SA) and Search-based Software Engineering (SBSE). The paper argues that these domains within Software Engineering should not work in isolation but exchange ideas among them. Such techniques can be particularly useful for situations which require learning from a lot of data or optimizing problems which need resources. This paper also presents resources and report our experience of tackling various SE problems.  Among the numerous questions that we try to answer are: (1) What are the various topics addressed by DSE?, (2) What are the types of data used by the researchers, (3) What research approaches do researchers use?, and  (4) What are levels of simplifications that can be added to these approaches. The paper briefly sets out to act as a practical guide to develop new data-driven techniques and also serve as a teaching resource.
\end{abstract}

@@

\keywords{}

@@

% \input{samplebody-conf}
\section{Introduction}
Software Analytics (SA) and Search-based Software Engineering (SBSE) are two of the newer fields in Software Engineering (SE) which deals with searching through a lot of data or options to find useful and actionable information. However, these two areas have evolved in isolation from each other. 
SA focuses on mining data generated during the software processes such as development, testing, etc. and uses various Machine Learning (ML) algorithms to extract information and relationships between different features of the collected data. 
Hidden in these data is precious and valuable information about the quality of software and services and the dynamics of software development. Mining of Software Repositories is a conference which is focused on software analytics. SBSE, on the other hand, involved formulating a software engineering problems as a search problem and uses meta-heuristic algorithms to balance competing constraints, trade-offs between concerns and requirements. SBSE has opened venues which promotes reformulation of various software engineering problems and has provided numerous inspirations for improving SE.  SBSE uses a data-driven approach to find the optimal (or near optimal) solutions to software engineering problems with the use of a fitness function which is used to differentiate between good (favorable) and bad (unfavorable) solutions. Symposium of Search-based Software Engineering (SSBSE) and Symposium of Search-based Software Testing (SBST) are the two venues which are dedicated to the dissemination of information in this area.

In a very loose sense, we can associate \textit{learning} with SA, since most of the work in this area attempts to learn from massive amounts of data, and \textit{optimization} with SBSE. In this paper, we show how learning and optimization are mutually inclusive---one cannot happen in isolation, which also goes to show that the field of SA and SBSE should not grow in isolation but rather exchange ideas.

Let us draw a comparison between these two fields. Please note that SA is very similar to traditional data mining but on software engineering data or data obtained from software processes.

\noindent\textbf{Data:} SBSE predominantly uses meta-heuristics algorithms such as NSGA-II~\cite{deb2000fast}, SPEA~\cite{zitzler2001spea2} to navigate the space of possible options. These algorithms rely on fitness functions to differentiate between good and bad solutions. This means there is no universal fitness function, for every problem, there is an associated fitness function. Similarly in SA, which uses different kinds of data miners to model the given data (or a problem). Different kinds of data miners work best of different kinds of data. This goes to show both techniques of SA and SBSE needs to be adjusted to fit the needs of the problem or the practitioner.

\noindent\textit{Design Choice:} The choice of the fitness function (in SBSE) and data miner (in SA) depends on the type of data used for that experiments.

\noindent\textbf{Goals:} In SBSE, the meta-heuristic algorithms used in an experiment depends on the number of goals of the experiment. A single goal or single-objective problem can be solved using algorithms such as \textit{Simulated Annealing}~\cite{van1987simulated} or hill climbing~\cite{rudlof1997stochastic}, whereas for multiple goal or multi-objective problems can be solved using algorithms such as NSGA-II, SPEA, MOEA/D~\cite{zhang2007moea}. In SA, if the data table has no goal columns, then this is an \textit{unsupervised learning problem} which can be addressed by (say) finding clusters of similar rows using, say, K-means or expectation maximization. An alternate approach, taken by the Apriori association rule learner, is to assume that every column is a goal and to look for what combinations of any values predict for any combination of any other. If a table has one goal, then this is a \textit{supervised learning problem} where the task is to find combinations of values from the other columns that predict for the goal values. Note that for datasets with one discrete goal feature, it is common to call that goal the class of the dataset. If a table has multiple goals, then the ML model (or combination of models) should trade-off in goal space by domination predicates (dominance counts, dominance depth, dominance ranks).

\noindent\textit{Design Choice:} The meta-heuristic algorithm (SBSE) or the data miner (SA) used for the experiments must be adjusted based on the number of goals in the experiment.

\noindent\textbf{Computational Overhead:} In SBSE, the size of the search space and the number of constraints in it, govern the computational budget of the experiment. Larger the search space or a highly constrained space means the meta-heuristic algorithms requires long run times or high computational budget. In SA, the size of the dataset dictates the time required to train the chosen ML algorithm, and hence various engineering decisions need to be taken to handle the size of the data.

\noindent\textit{Design Choice:} The computational budget required to run both SBSE and SA techniques are governed by the size of the search space and data respectively.

\noindent\textbf{Parameter Tuning:} In SBSE, the convergence of the meta-heuristic algorithms are governed by the parameters of the meta-heuristic algorithm~\cite{eiben2011parameter}. Similarly, in SA, the effectiveness of the ML algorithm used is controlled by its various parameters~\cite{fu2016tuning, fu2016differential, tantithamthavorn2016automated}.

\noindent\textit{Design Choice:} The performance of the meta-heuristic algorithm (in SBSE), and machine learning algorithms (in SA) is governed by the parameters used in them.

So, syntactically, there is much overlap between design choices in SBSE and SA. Hence we can say that learning and optimization go hand in hand and we should not treat both of these techniques in isolation. Instead, we should exchange ideas between these two domains. We call the marriage of these two fields by Data-Driven Search-based Software Engineering (DSE\footnote{DDSSE might be a more appropriate abbreviation but let us remove redundancy in the name.}). We appeal to the community that it is time to modify our approaches. It is not longer SBSE or SA. Let us tie these two domain and exchange ideas to bring the learning and searching closer---thereby learning from each other. \footnote{The geneses of the ideas presented in this paper happened in the \href{http://shonan.nii.ac.jp/shonan/blog/2016/09/08/data-driven-search-based-software-engineering/}{No.105 Data-Driven Search-Based Software Engineering}, where leading practitioners came together to propose and exchange ideas.}

The contributions of the paper are:
\begin{itemize}
    \item To show that optimization (SBSE) and learning (SA) goes hand in hand,
    \item Provide resources to seed various research,
    \item Provide teaching resources, which can be used to create DSE courses, and
    \item Based on our experience, various strategies which can be used to make these DSE techniques more efficient.
\end{itemize}

\section{Background}

\subsection{Search-based Software Engineering}
The term search-based software engineering was coined by Jones and Harman~\cite{harman2001search} in 2001 and over the years have been used in various fields of software engineering for example, requirements~\cite{ZhangHL13, chen2017beyond}, automatic program repair~\cite{le2012genprog}, Software Product Lines~\cite{chen2017sampling, sayyad2013value, guo2017smtibea}, Performance configuration optimization~\cite{nair2017using, guo2017data, oh2017finding, nair2018finding} to name of few. SBSE has been applied to other fields and has their own surveys such as design~\cite{raiha2010survey}, model-driven engineering~\cite{boussaid2017survey}, genetic improvement of programs~\cite{petke2017genetic}, refactoring~\cite{mariani2017systematic}, Testing~\cite{silva2017systematic, khari2017extensive} as well as more general surveys~\cite{clarke2003reformulating, harman2007current}. To convert a software engineering problem to an optimization problem, most prior works use the following steps.

\noindent\textbf{Data Collection or Model building: } SBSE process can either rely on a model, which represent a software process~\cite{boehm1995cost} or can be directly applied to any software engineering problem including problems which requires evaluating a solution by running a specific benchmark~\cite{krall2015gale}. This is very similar to model-driven engineering (MDE), and there are several works drawing parallels of these two fields~\cite{boussaid2017survey, kessentini2013searching}. Along with the data collection, it is also essential to determine the performance metric which needs to be optimized (the goal of optimization).

\noindent\textbf{Representation: } The process of search starts by creating random solutions (valid or invalid) of certain problems. In some cases, practitioners have also tried to seed the initial population for faster convergence of the search process~\cite{saber2017seeding, chen2017beyond, chen2017sampling, henard2015combining}. The representation of these solutions is significant to generate the landscape of the problem (details in Section~\ref{sec:help}). The general representation of the problem uses either boolean or numerics. This representation can also be called as the \textit{Decision Space}.

\noindent\textbf{Fitness Function: } A fitness function maps the solution (which is represented using numerics) to a numeric scale (also called as \textit{Objective Space}) which is used to distinguish between good and not so good solutions. This measure is a domain-specific measure (single objective) or measures (multi-objective) which is useful for the practitioners. Examples of various fitness used in SBSE are (i) difference between the lowest
and the highest number of transformation elements~\cite{fleck2017model} or (ii) number of test cases passed~\cite{oliveira2016improved}. The fitness function determines the fitness landscape of the problem---which characterizes the problem and defines the `hardness' of the problem. Fitness landscape should not be too flat, which means the meta-heuristic algorithms cannot find the best possible direction to explore. Simply put fitness function is a transformation function which converts the point in the decision space to an objective space.

\noindent\textbf{Operators: } Meta-heuristic algorithms require \textit{recombination operators} such as crossover and mutation operators which can be applied along with \textit{elitism operator} to simulate the `survival of the fittest' strategy. The combination operator combines the features of various solutions in the population to generate new candidates---effectiveness of which is then measured using the fitness function. The elitism operator uses these fitness values to eliminate the not so good solutions from the population. The intuition behind this process is to remove bad solutions and move towards better solutions.

\subsection{Software Analytics}\label{sec:SA}
Most decisions (resource allocation, effort estimation) in software engineering are based on the perception of people (decision makers). As the size of the software systems and the number of interactions (developers and consumers) grow, there is a growing need for data-driven decision-making solutions. Additionally, decision-makers often struggle to come to a conclusion or common consensus (when more than one decision maker is involved) under a lot of uncertainty. They constantly look for techniques that return solutions which are reliable and can assist the decision-making processes which in turn affect the product, team, and the processes. These concerns have drawn much attention to software measurement, software quality, and software cost/ effort estimation, namely descriptive and predictive analytics~\cite{bener2015lessons}.

SA can be used by the practitioners to build reliable models based on historical data or empirical observations. Using these models, decision makers can make informed decisions about how to manage resources, cost, developers, etc. Most of the work in SA follows the following steps.

\noindent\textbf{Data Collection:} The primary process of SA starts with the collection of data from the software processes. This can be CK metrics from the source code~\cite{subramanyam2003empirical}, feature models from software product lines~\cite{sayyad2013value}, text dumps from Stackoverflow~\cite{agrawalwrong} to name a few. These data are divided into columns or features~\footnote{In case of textual data, it has to be tokenized and converted into a table.} and rows. Each column represents the features of the data whereas each row represents a data point (modules in case of CK metrics). Please note, the data may or may not have a target variable---used to quantify the performance of the data point. For example, CK metrics have been used to predict for defects in a software module, and in this case, the target variable associated to the CK metrics of a specific module can be the number of defects found in that module. Another aspect can be the preparation of data---fixing the weaker regions of the
training data using \textit{resampling techniques}, so that the performance of the model trained using the (resampled) data improves dramatically~\cite{agrawal2017better, bennin2017mahakil}. It should be noted that in this scenario, it is assumed that there is some historical data (from past projects) which can be used to find the target variable (associated number of defects in our setting). But, there are instances, where such historical data is unavailable, which means the data (CK metrics) does not have a corresponding target variable~\cite{yang2016effort}.

\noindent\textbf{Clustering:} The data can be \textit{clustered} into various sub-classes either based on the target variable or based on the feature space. This is done to overcome the effects of data heterogeneity---the data can have local regions which behave very differently than other (sub-)regions. Prior work in this area has found that treating these regions in isolation can result in inconclusive results~\cite{menzies2011local, menzies2013local, bettenburg2012think}. A more profound finding by the prior work is that practitioners should not seek to find general principals which can be applied to many projects but rather focus on ways to find the best local lessons for groups of related projects.

\noindent\textbf{Pruning: }Data Pruning can be divided into two major classes---(i) Row Pruning, and (ii) Columns Pruning. Row pruning can be either done to divide the available data into different subsets (similar to clustering) or to remove columns which adds uncertainty or noise to the data. Column Pruning is beneficial only to consider columns which affect the target variable. This removes unwanted noise from the data as well as reduces the cost of data collection in the future projects~\cite{kirsopp2003case, hall2003benchmarking, chen2005finding, gao2011choosing}.\footnote{If a model is built using all the features available; it means that all the future projects need to extract all these features from the available data. This increases the cost of data collection in the future. } Other advantages of data pruning are: reducing variance in model predictions, removing irrelevant data points and noise removal.

\noindent\textbf{Learner selection and Parameter Tuning: } The machine learning algorithms or the learner used in SA experiments are selected carefully. ML domain has several techniques and best practices that are used to choose the best algorithms~\cite{witten2016data, malkomes2016bayesian}. Hall et al. showed a ranking list of various learners, which best suited for defect prediction in software projects. However, these ranking cannot be trusted without careful varying the parameters of the model. The ranking of the learners, as presented by Hall et al.~\cite{hall2012systematic}, change drastically when the parameters of the learners are changed~\cite{fu2016tuning, tantithamthavorn2016automated, fu2017easy}.

\section{Data-driven Software Engineering: How SBSE met SA}
A lot of work in SBSE and SA have tried and solved various software engineering problems using very different techniques. SBSE focused on formalizing the problem as an optimization problem, where as SA treated the problem as a learning problem---building a model which can be used to characterize the problem or feature space. Even though the strategies used in both these domains are different, we claim that practitioner in both the domains can learn from the advances in the other community. In this section, we would describe simple scenarios on how techniques from SA can be used to increase the effectiveness of SBSE techniques and vice versa.

\begin{figure}[t]
\small
\hspace{0.4cm}\begin{lstlisting}[xrightmargin=5.0ex,mathescape,frame=none,numbers=right]
  # A Traditional bare-bone algorithm
  def EA(problem, pop_size, budget):
  # Generate the decision space by randomly generating solutions
  initial_pop = [generate(random=True) for _ in range(pop_size)]
  # Evaluate all the individuals in the initial_pop using problem 
  # specific fitness function
  pop = [individual.fitness(problem) for individual in initial_pop]
  while budget > 0:
    # Generate mutant or new individuals by recombining individuals from pop
    # Size of new_pop is equal to size of pop.
    new_pop = recombination_op(pop)
    # Evaluate the mutants from the new_pop
    new_pop = [mutant.fitness(problem) for mutant in new_pop]
    # Select the best performing individual from pop + new_pop
    pop = elitism(new_pop + pop)
    # Reduce the budget by 1. 
    budget -= 1
  return pop

\end{lstlisting}
\caption{\small{Psuedocode of Evolutionary Algorithm.}
}
\label{fig:EA}  
\end{figure}

\subsection{How techniques from SA can help SBSE?}\label{sec:help}

The whole process of SBSE can be simplified using a \textit{black-box model}---where a searcher (meta-heuristic algorithm) probes the black-box (fitness evaluation) to find the maximum or minimum value. The process can also be simplified using the notion of \textit{landscape}. A problem has two landscape associated with it namely (i) decision landscape, and (ii) objective or fitness landscape. It is fair to assume that these landscapes are correlated, which means that data points which are close in decision space are also close in the objective space. 
The objective of a searcher is to efficiently search through the decision space and probe the black-box model to the maximum (or minimum) in the fitness landscape.

In essence, the whole process of searching can be described as a strategy to maneuver the fitness landscape. Using this terminology, let us try to redefine a meta-heuristic based on the black-box model. We use an Evolutionary algorithm as a representative meta-heuristic algorithm since EA has by far the most widely used algorithm in practise~\cite{nair2016accidental}.
An Evolutionary Algorithm (EA) uses a population-based method to find the maximum or minimum value in the function (or a fitness landscape). It is called evolutionary since it iteratively selects `promising' solutions and improves them by using recombination operators. It starts by \textit{generating} a population (either randomly or using some heuristics or checks). These are solutions evaluated using fitness function or probing the black-box. The combination operators use the population and the associated fitness function to generate a new population. This new population ideally explores areas which are considered more `promising'---higher fitness value than the prior generation. This means EAs uses the current population to move towards the most promising region or regions of the fitness space.

Since SBSE, depends on a population-based strategy, it is plagued with slow runtimes and convergence rates. This makes it infeasible for cases where each model evaluation is expensive. For example, if we want to find the (near) optimal configuration (e.g., minimizing the load time for a specific benchmark) of an Apache web server. The configuration space in our setting represents the decision space with constraints---constrained by the feature model, which represents the constraints among the configuration options. The fitness function represents executing the workload for a specific configuration. This can be very time consuming, and SBSE techniques is not a viable option for these kinds of problems. In the optimization literature, these types of problems are referred to as \textit{expensive problems}.

In SA, an ML algorithm is used to model the data. The model is then used to answer questions of the stakeholder (as seen in section~\ref{sec:SA}). In problems, where fitness evaluation is expensive, we can use an ML model to learn the (approximate) landscape of the configuration space. The machine learning model now becomes a cheap black-box, which can be probed for free (since predicting a data point (using an ML model) requires almost no time). This is also known as surrogate-based optimization (refer to \cite{jin2011surrogate} for more details).

In Figure~\ref{fig:EA},  we list a generic algorithm for EA. The algorithm starts with randomly generating a population (\textit{initial\_pop}) of size (\textit{pop\_size}) (Line 4). All the individuals in the \textit{initial\_pop} is evaluated and stored in \textit{pop} (Line 5). In literature, the size of the \textit{new\_pop} is equal to the size of the \textit{pop}. A new population (\textit{new\_pop}) is generated using recombination operators (Line 11). The fitness of the newly generated individuals are evaluated (Line 13) after which an elitism operator is used to filter out the not so `promising' individuals (Line 15).  This process continues till the budget runs out (number of generations in our setting).
EA can be slow since it evaluates all the newly created individuals in the while loop (Line 8-17) and makes it unsuitable to use for problems where the cost (time or resources) required to evaluate the \textit{fitness} of an individual is high.

\begin{figure}[t]
\small
\hspace{0.4cm}
\begin{lstlisting}[xleftmargin=5.0ex,mathescape,frame=none,numbers=left,linebackgroundcolor={%
        \ifnum\value{lstnumber}>11
            \ifnum\value{lstnumber}<25
                \color{gray!25}
            \fi
        \fi},autogobble=true,]
  # A SA based EA
  def EA(problem, pop_size, budget, gen_budget):
  # List to collect all evaluated individuals
  evaluated = list()
  # Generate the decision space by randomly generating solutions
  initial_pop = [generate(random=True) for _ in range(pop_size)]
  # Evaluate all the individuals in the initial_pop using problem 
  # specific fitness function
  pop = [individual.fitness(problem) for individual in initial_pop]
  evaluated += pop
  while budget > 0:
    # Generate mutant or new individuals by recombining individuals from pop.
    # Size of new_pop is much larger to size of pop.
    new_pop = recombination_op(pop){
    # Train a machine learning model using individuals from pop 
    model = ML.train(evaluated)
    # Evaluate new_pop using the model---cheaper than individual.fitness 
    pred_new_pop = [model.fitness(individual) for individual in new_pop]
    # Only evaluate the most promising candidates defined by gen_budget
    new_pop = sorted(pred_new_pop, reverse=True).select(gen_budget)
    # Evaluate the individual which are predicted to be promising
    new_pop = [selected.fitness(problem) for selected in new_pop]
    # Adding evaluated individuals
    evaluated += new_pop
    # Select the best performing individual from pop + new_pop
    pop = elitism(pred_new_pop + pop)
    # Reduce the budget by 1. 
    budget -= 1
  return pop

\end{lstlisting}
\caption{\small{Psuedocode of combining SA technique to a SBSE technique.}
}
\label{fig:SAEA}  
\end{figure}

To overcome this problem of EA, we can use an ML algorithm to learn the landscape of the problem. We can use already evaluated individuals to train an ML model to learn the landscape and only evaluate the promising individuals (as predicted by the ML model). In Figure~\ref{fig:SAEA}, we show the EA algorithm which incorporates an ML model (shown in the shaded region)  to reduce the number of evaluations. Similar to EA, a \textit{new\_pop} is generated by recombining the individuals from the \textit{pop} (Line 11). The only difference (from EA) is, the size of the \textit{new\_pop} is much larger than the size \textit{pop}. The already evaluated individuals are used to train an ML model (Line 13). This trained model, \textit{model}, is then used to predict the fitness values of the solutions in \textit{new\_pop} (Line 18). The \textit{pred\_new\_pop} is then sorted (assuming larger fitness value is good) and only \textit{gen\_budget} number of individuals are selected (Line 22). All the individuals in \textit{gen\_budget} is stored in the \textit{evaluated} list for subsequent model training (Line 24). This process continues till a budget is reached.

This is a very simplistic approach to reduce the cost of running the EAs and increasing the convergence rate. For example, in the Figures~\ref{fig:EA},~\ref{fig:SAEA}, if the $\mathit{pop\_size}=100$, $\mathit{budget}=10$, and $\mathit{gen\_budget}=10$, the total number of fitness evaluations (probing) done by EA is 1100 ($100 + 10\times100$), whereas in SA inspired EA it is 200 ($100 + 10\times10$). Please note, this is only one of the many possible strategies to use an SA technique to improve the performance of an EA (both in terms of time and convergence). In section~\ref{sec:scenarios}, we describe several strategies to solve an SBSE problem using techniques from SA.

\begin{figure}[t]
\small
\hspace{0.4cm}
\begin{lstlisting}[xrightmargin=5.0ex,mathescape,frame=none,numbers=right]
  # Analytic Task
  def model_building(ml_alg, train_data, test_data,parameter=default):
    # Train the classifier using the default parameter using the training data
    model = ml_alg(parameter).train(train_data.features, train_data.label)
    # Trained model is used to predict the labels of the test data
    return model.predict(test_data.features)
\end{lstlisting}
\caption{\small{Psuedocode of the analytic task}
}
\label{fig:SA_default}  
\end{figure}

\begin{figure}[t]
\small
\hspace{0.4cm}
\begin{lstlisting}[xrightmargin=5.0ex,mathescape,frame=none,numbers=right,linebackgroundcolor={%
        \ifnum\value{lstnumber}>9
            \ifnum\value{lstnumber}<30
                \color{gray!25}
            \fi
        \fi},autogobble=true,]
  # Analytic Task
  def model_building(ml_alg, parameter_space, train_data, test_data):
    # Find the best parameter to train a classifier using the training data
    parameter = find_best_model(ml_alg, parameter, space, train_data)
    # Train a model using the parameter 
    model = ml_alg(parameter).train(train_data.features, train_data.label)
    # Trained model is used to predict the labels of the test data
    return model.predict(test_data.features)
  # An EA
  def find_best_model(ml_alg, parameter_space, train_data, budget):
    # The training data is split into two. 
    train_1, train_2 = train_data.split()
    # Initial population---parameters of classifier is generated
    initial_pop = [generate(parameter_space) for _ in range(pop_size)]
    # The performance of the parameters is measured by training a model with a 
    # specific parameter on the train_1 and testing on train_2
    pop = [ind.fitness(ml_alg,train_1,train_2) for ind in initial_pop]
    while budget > 0:
      # Generate mutant or new individuals by recombining individuals from pop
      # Size of new_pop is equal to size of pop.
      new_pop = recombination_op(pop)
      # Evaluate the mutants from the new_pop
      new_pop = [mut.fitness(ml_alg,train_1,train_2) for mut in new_pop]
      # Select the best performing individual from pop + new_pop
      pop = elitism(new_pop + pop, performance)
      # Reduce the budget by 1. 
      budget -= 1
    # Return the parameter which gave the best result
    return max(pop)

\end{lstlisting}
\caption{\small{Psuedocode of combining SBSE technique to enhance SA.}
}
\label{fig:EASA}  
\end{figure}

\subsection{How strategies from SBSE can help SA}

SA utilizes data-driven approaches to enable software practitioners to perform data exploration and analysis to obtain insightful and actionable information. To handle the volume of the data produced during the software development problem, SA practitioners use analytic tools to find patterns in the data. The process of SA can be broadly divided into the data preparation and the analytic task which is performed on the data. An example of SA task can be a classification task, such as defect prediction, where historical data is used to classify the software modules as defect prone or non-defective. This can be then used to concentrate testing efforts.

In this section, we would concentrate on how techniques from SBSE can be used to improve the analytic task of software analytics. Once the data is prepared, cleaned, and resampled for the practitioner to use, the next important task is to find the best suited analytic tools. The community is divided as to which analytic tool is the most effective. There have been several ranking studies, which tried to rank classifier based on their effectiveness~\cite{lessmann2008benchmarking,hall2012systematic,elish2008predicting,menzies2010defect,gondra2008applying}. However, they do not consider the parameter of the classifiers. In the machine learning community, hyper-parameter tuning is one of the most critical steps of the model building---which was not explored, until recently, in SA. Hyper-parameter tuning requires adjusting the parameters (of classifiers) to maximize the performance measure, such as Precision, recall. This problem in itself is an optimization problem, which can be solved using techniques from SBSE.

In Figure~\ref{fig:SA_default}, we list a generic algorithm for an SA task. The algorithms start with training a classifier (\textit{ml\_alg}) using the training data or the historical data (Line 4). The parameter used to train the classifier is the default parameter---as provided by the library developers. The trained classifier (\textit{model}) is then used to predict the labels of the testing data (Line 6). Testing data in our setting, are the code base which is yet to be tested.

In Figure~\ref{fig:EASA}, we list a generic algorithm of how the effectiveness of an SA task can be enhanced by using the SBSE technique. The code highlighted in \textcolor{gray}{gray} is the SBSE technique used. Here, unlike the usual SA task, rather than using the default parameter, the practitioner provides the parameter space---space which contains the best parameter (Line 2). Please note that the size of the parameter space is decided based on the budget (\textit{budget}) available to perform hyper-parameter tuning. Next step is to used the EA (\textit{find\_best\_model}) to find the parameters which would result in the best classifier (\textit{model}) (Line 4). In \textit{find\_best\_model}, the training data (\textit{train\_data}) is divided into two parts (\textit{train\_1} and \textit{train\_2}) (Line 12). The initial population (\textit{initial\_pop}) is generated from within the parameter space provided by the practitioner (Line 14). The initial population contains parameters of the classifier. The fitness of the parameter is evaluated by training  the classifier (\textit{ml\_alg}) using \textit{train\_1} and testing its performance on \textit{train\_2} (Line 17). The search process to find the best parameter is same as described in Figure~\ref{fig:EA}. Once EA terminates the best parameter is selected and returned (Line 29). The model is then trained using the parameter returned by \textit{find\_best\_model} (Line 6). The trained classifier (\textit{model}) is then used to predict the labels of the testing data (Line 8). This process has proven to be very useful in multiple domains~\cite{fu2016tuning, fu2016differential, fu2017easy, agrawal2017better, tantithamthavorn2016automated}.

\cite{mathew2017shorter}

\begin{table*}[]
\centering
